---
title: "Final Project Milestone 4"
author: "Michael Wu"
date: "3/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reprex)
library(janitor)
library(gt)
library(readr)
library(readxl)
library(rvest)
library(ggthemes)
```

### the Coronavirus Project.

With the WHO classifying the risk of the novel coronavirus to its highest levels today and public health magnates like Bill Gates calling the disease a pandemic, COVID-19 is both literally and figuratively unlike anything we have ever seen before.

The public response around the globe has been that of sheer panic, with ramifications of the disease rippling through all facets of society. Grocery store shelves in Italy have gone empty from the public’s fear of food shortages and quarantine, investors on Wall Street have been whipped into a near state of hysteria with the DJI losing 10% of its value in just six sessions – the quickest correction ever.

For my Gov 1005 final project, I thought it would useful for the public discussion about the virus to explore how COVID-19 has affected global transportation chains compared to previous epidemics like the SARS outbreak of 2002, the H1N1 pandemic of 2009, and the Ebola outbreak from 2014-2016. Perhaps, seeing the global response in terms of executing travel bans or other forms of travel restrictions on previous outbreaks and how successful we were in stopping the spread of those outbreaks will provide a useful lesson for countries caught in today’s outbreak of when it is appropriate to institute severe travel restrictions. It’s probably also important that I standardize these spread results with some discussion related to a r-naught value (how spreadable a certain disease is), which I will do at a later point in time.

As of this third milestone, I have managed to successfully obtain, clean, and tidy data on the coronavirus’s spread from Johns Hopkins University’s GitHub page. A selected sample of this data is displayed below in table format. I am still working on obtaining the current and historic flight data from flightradar24 – I signed up for a seven-day free trail account and sent them a request for this data last night and they have yet to respond back to me. I have also obtained some data from the SARS outbreak in 2002. There remains however, some additional work that I need to do, which is conduct some web scraping from the url links provided in the .xlsx file which has the breakdown of the total number of infections/deaths by country. I’m still searching for good Ebola and H1N1 data sets, but am confident that I will find these in due time.

As of the fourth milestone, I have created two new plots of the case counts in iran and South korea and begun preliminarily cleaning the data sets. I have also updated the data sets so they are up to date from this week's numbers. I came up with a really interesting new lead to build upon this data that I will also discuss below, and was wondering if you had any potential pointers on where I could find these data sets.

I just came across an extremely disturbing report that the Iranian government is systematically suppressing nurses/doctors into changing their coronavirus death counts by changing the cause of death.

After panicking for a few minutes about this and telling my friends about it, I realized that it would make for a really interesting final project for me to explore the current number of deaths from Iran over the past two weeks of the outbreak (maybe from hospital data or the country's reporting?), cross-reference that with what the UN projects their death rate/population to be, subtract out the number of reported cases, and see if I can find any number of unaccounted for deaths.

I've found some UN population projection data here,
https://www.un.org/en/development/desa/population/publications/database/index.asp

But I'm having trouble finding the number of deaths out of Iran. Do you have any advice or thoughts on where I could find this?

A secondary lead that I could follow is that I came across some comments on reddit which made a lot of sense to me and sparked a new idea. The comment was that a lot of these coronavirus cases that are not being reported/tested are likely being reported as seasonal flu deaths. So, I think it would make a lot of sense to somehow plot the number of cases/fatalities of the seasonal flu during this flu season, and see if there appear to be any unnatural spikes or unexplainable aspects of the data. I've found some data on influenza from the WHO but they seem to have a really small sample size. Any advice?
https://www.who.int/influenza/gisrs_laboratory/flunet/en/


### New Updates for the Fifth Milestone

Hello! It's been a bit of crazy week having to move back home from campus and get adjusted to the quarantine life, but I feel reasonably satisfied with my progress as of the fifth milestone. First, I have updated the datasets so that they are in line with the most updated realtime data from the John Hopkins GitHub page. I have also created three separate graphs that plot the case numbers in the most affected countries and also the most affected provinces and states within China and US (wherever there's province/state data).

Something really simple that I think would be interesting is to somehow also factor in each country's overall population so we can see if larger provinces, states, or countries experience more rapid case growth. I'm still working on obtaining the previous flu data from the WHO and will try to get that up and running next week. Please let me know if you have any general pieces of feedback!

$~$

### Preliminary Data Sets for the Coronavirus and SARS Outbreaks

```{r loading_data, echo=FALSE, results='hide'}
# Loading the untidy version of the coronavirus data forked from the Johns Hopkins' GitHub repo
# Specifying column types to avoid any forms of errors

untidy_covidata <- read_csv('JH-covid19-rawdata/time_series_19-covid-Confirmed.csv', col_types = cols(
                              .default = col_double(),
                              `Province/State` = col_character(),
                              `Country/Region` = col_character()
                              )
                            ) %>%
  clean_names()

# This read_excel output is giving me some weird output of "new names: * `` -> ...1" and I am unsure of how to fix it
# UPDATE: was able to fix it with help from the following rstudio community post
# https://community.rstudio.com/t/how-to-make-read-excel-stops-messaging-when-using-col-names-false/30495/4
# Loaded from Zhong Tron's Kaggle page, scraped originally from the WHO

sarsdata <- read_excel(path = 'sars-rawdata/sars_final.xlsx', .name_repair = "minimal") %>%
  clean_names()
```

```{r tidying-data, echo=FALSE, results='hide'}
# Tidying the untidy_covidata tibble using pivot_longer
# Pivoting so that dates will now be row data entries instead of columns as they were originally

covidata <- untidy_covidata %>%
  pivot_longer(
    cols = starts_with("x"),
    names_to = "date",
    names_prefix = "x",
    values_to = "cases",
    values_drop_na = TRUE
  )
```

$~$

```{r display-untidy-table, echo=FALSE}
# Creating a gt() plot of the untidy_covidata just for reference
# Including two important footnotes for this table
# Only want to select a few columns that represent dates just to give the viewers a taste of the data

untidy_covidata %>%
  slice(1:5) %>%
  select(province_state, country_region, lat, long, x1_22_20, x1_23_20, x1_24_20, x1_25_20, x2_27_20) %>%
  gt() %>% 
  tab_header(
    title = "Glimpse of Untidy Covid-19 Data"
  ) %>%
  cols_label(
    province_state = "Province/State",
    country_region = "Country",
    lat = "Latitude",
    long = "Longitude"
  ) %>%
  cols_align(
    align = "center"
  ) %>%
  tab_footnote(
    footnote = "Source: Johns Hopkins University, Ryan Lau on Github",
    locations = cells_title(
      groups = "title"
    )
  ) %>%
  tab_footnote(
    footnote = "First five rows displayed only",
    locations = cells_title(
      groups = "title"
    )
  )
```

$~$

```{r display-tidy-table, echo=FALSE}
# Creating a gt() plot of the tidied covidata variable
# Nothing too groundbreaking with the code here, just a standard gt() tibble

covidata %>%
  slice(1:5)%>%
  gt() %>% 
  tab_header(
    title = "Glimpse of Tidied Covid-19 Data"
  ) %>%
  cols_label(
    province_state = "Province/State",
    country_region = "Country",
    lat = "Latitude",
    long = "Longitude",
    date = "Date",
    cases = "Number of Cases"
  ) %>%
  cols_align(
    align = "center"
  ) %>%
  tab_footnote(
    footnote = "Source: Johns Hopkins University, Ryan Lau on Github",
    locations = cells_title(
      groups = "title"
    )
  ) %>%
  tab_footnote(
    footnote = "First five rows displayed only",
    locations = cells_title(
      groups = "title"
    )
  )
  
```

$~$

```{r plot-sk-covidata, echo = FALSE}
sk_covidata <- covidata %>%
  filter(country_region == "Korea, South")

sk_plot <- ggplot(sk_covidata, aes(x = as.Date(date,"%m_%d_%y"), y = cases)) +
  geom_line() +
  labs(title = "South Korea Covid-19 Cases",
       subtitle = "As of March 20th, 2020",
       x = "Case Count",
       y = "Date") +
  theme_classic()
  
```

$~$

```{r plot-iran-covidata, echo = FALSE}
iran_covidata <- covidata %>%
  filter(country_region == "Iran")

iran_plot <- ggplot(iran_covidata, aes(x = as.Date(date,"%m_%d_%y"), y = cases)) +
  geom_line() +
  labs(title = "Iran Covid-19 Cases",
       subtitle = "As of March 20th, 2020",
       x = "Case Count",
       y = "Date") +
  theme_classic()
  
```

```{r plot-italy-covidata, echo = FALSE}
italy_covidata <- covidata %>%
  filter(country_region == "Italy")

italy_plot <- ggplot(italy_covidata, aes(x = as.Date(date,"%m_%d_%y"), y = cases)) +
  geom_line() +
  labs(title = "Italy Covid-19 Cases",
       subtitle = "As of March 20th, 2020",
       x = "Case Count",
       y = "Date") +
  theme_classic()
  
```

```{r plot-country-covidata, echo = FALSE}
filtered_covidata <- covidata %>%
  filter(country_region %in% c("Italy", "Korea, South", "Spain", "France", "Germany", "Iran")) %>%
  filter(is.na(province_state) | province_state == "France")

ggplot(filtered_covidata, aes(x = as.Date(date,"%m_%d_%y"), y = cases, color = country_region)) +
  geom_line() +
  labs(title = "Confirmed Covid-19 Cases of the Most Affected Countries",
       subtitle = "As of March 20th, Excluding China and the United States",
       x = "Date",
       y = "Case Count") +
  theme_classic()
  
```

```{r plot-US-covidata, echo = FALSE}
# Find the most affected states in the United States
# Would use log scales like in China's case below, but the case reporting is so erratic
# that the graphs for log scales on the y-axis looks so poor.

US_covidata <- covidata %>%
  filter(country_region == "US") %>%
  filter(province_state %in% c("New York", "Washington", "California", "New Jersey"))

ggplot(US_covidata, aes(x = as.Date(date,"%m_%d_%y"), y = cases, color = province_state)) +
  geom_line() +
  labs(title = "Confirmed Covid-19 Cases in China's 4 Most Affected Provinces",
       subtitle = "As of March 20th, In Log Scale",
       x = "Date",
       y = "Case Count") +
  theme_classic()
  
```

```{r plot-china-covidata, echo = FALSE}
# Find the most affected provinces in China
# Need to use log-scales because Hubei has so more many cases than all other provinces combined

China_covidata <- covidata %>%
  filter(country_region == "China") %>%
  filter(province_state %in% c("Hubei", "Guangdong", "Henan", "Zhejiang"))

ggplot(China_covidata, aes(x = as.Date(date,"%m_%d_%y"), y = cases, color = province_state)) +
  geom_line() +
  labs(title = "Confirmed Covid-19 Cases in China's 4 Most Affected Provinces",
       subtitle = "In Log Scale",
       x = "Date",
       y = "Case Count") +
  theme_classic() +
  scale_y_continuous(trans='log10')
  
```

$~$

```{r display-sars-table, echo=FALSE}
# Creating a gt() plot of the tidied csars tibble
# Nothing too groundbreaking with the code here, just a standard gt() tibble

sarsdata %>%
  select(date, infected, mortality) %>%
  slice(1:5) %>%
  gt() %>% 
  tab_header(
    title = "Glimpse of SARS Data"
  ) %>%
  cols_label(
    date = "Date",
    infected = "Infected",
    mortality = "Deaths"
  ) %>%
  cols_align(
    align = "center"
  ) %>%
  tab_footnote(
    footnote = "Source: World Health Organization, Zhong Tron on Kaggle",
    locations = cells_title(
      groups = "title"
    )
  ) %>%
  tab_footnote(
    footnote = "First five rows displayed only",
    locations = cells_title(
      groups = "title"
    )
  )
  
```

$~$