---
title: "Final Project Milestone 5"
author: "Michael Wu"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(reprex)
library(janitor)
library(gt)
library(readr)
library(readxl)
library(rvest)
library(ggthemes)
```

### the Coronavirus Project.

With the WHO classifying the risk of the novel coronavirus as a pandemic, COVID-19 is both literally and figuratively unlike anything we have ever seen before.

The public response around the globe has been that of sheer panic, with ramifications of the disease rippling through all facets of society. Grocery store shelves in Italy have gone empty from the public’s fear of food shortages and quarantine, investors on Wall Street have been whipped into a near state of hysteria with the DJI losing 35% of its value in just three weeks – one of the quickest corrections ever.


```{r loading_data, echo=FALSE, results='hide'}
# Loading the untidy version of the coronavirus data forked from the Johns Hopkins' GitHub repo
# Specifying column types to avoid any forms of errors

untidy_covidata <- read_csv('JH-covid19-rawdata/time_series_19-covid-Confirmed.csv', col_types = cols(
                              .default = col_double(),
                              `Province/State` = col_character(),
                              `Country/Region` = col_character()
                              )
                            ) %>%
  clean_names()

# This read_excel output is giving me some weird output of "new names: * `` -> ...1" and I am unsure of how to fix it
# UPDATE: was able to fix it with help from the following rstudio community post
# https://community.rstudio.com/t/how-to-make-read-excel-stops-messaging-when-using-col-names-false/30495/4
# Loaded from Zhong Tron's Kaggle page, scraped originally from the WHO

sarsdata <- read_excel(path = 'sars-rawdata/sars_final.xlsx', .name_repair = "minimal") %>%
  clean_names()
```

```{r tidying-data, echo=FALSE, fig.align='center'}
# Tidying the untidy_covidata tibble using pivot_longer
# Pivoting so that dates will now be row data entries instead of columns as they were originally

covidata <- untidy_covidata %>%
  pivot_longer(
    cols = starts_with("x"),
    names_to = "date",
    names_prefix = "x",
    values_to = "cases",
    values_drop_na = TRUE
  )

filtered_covidata <- covidata %>%
  filter(country_region %in% c("Italy", "Korea, South", "Spain", "France", "Germany", "Iran")) %>%
  filter(is.na(province_state) | province_state == "France")

ggplot(filtered_covidata, aes(x = as.Date(date,"%m_%d_%y"), y = cases, color = country_region)) +
  geom_line() +
  labs(title = "Confirmed Covid-19 Cases of the Most Affected Countries",
       subtitle = "As of March 20th, Excluding China and the United States",
       x = "Date",
       y = "Case Count") +
  theme_classic()
  
```



$~$
